```{r setup, message=FALSE, warning=FALSE}
source("./0_Libraries.R")
```

# Study area maps

```{r study-maps-pkgs, message=FALSE, warning=FALSE}
library(tidycensus)
library(tidyterra)
library(ggnewscale)
library(cowplot)
library(scales)
library(ggspatial)
```

```{r study-maps-data, message=FALSE, warning=FALSE}
la_city   <- get_city_boundary(proj_crs)
la_county <- get_county_boundary(proj_crs)

# City boundary clipped to county — removes ocean portions of the outline
la_city_land <- st_intersection(la_city, la_county)

la_water_city <- area_water(state = proj_state, county = proj_county) %>%
  st_transform(proj_crs) %>%
  clipintersect_boundary(la_city, buffer = 0)

# California state — used to derive ocean polygon for elevation map
ca <- tigris::states(cb = TRUE, class = "sf", progress_bar = FALSE) %>%
  filter(STUSPS == "CA") %>%
  st_transform(proj_crs)

# View extent: 20 km buffer around LA City
city_view_bbox <- st_as_sfc(st_bbox(st_buffer(la_city, 20000))) %>% st_set_crs(proj_crs)

# Ocean = view extent minus CA land (overlaid on top of terrain to mask ocean DEM values)
ocean <- st_difference(city_view_bbox, st_union(ca))

```

## Map 1 — Population density by census tract

```{r map-pop-density, message=FALSE, warning=FALSE}
# ACS total population per tract
ct_geom <- get_census_tracts(base_path, proj_crs, proj_state, proj_year, proj_county) %>%
  mutate(GEOID = as.character(GEOID))

ct_pop <- get_acs(
  geography = "tract",
  variables = "B01003_001",
  state     = proj_state,
  county    = "Los Angeles",
  year      = proj_year,
  progress  = FALSE
)

ct_density <- ct_geom %>%
  inner_join(ct_pop, by = "GEOID") %>%
  filter(lengths(st_intersects(., la_city)) > 0) %>%
  mutate(
    aland_km2   = ALAND / 1e6,
    pop_density = if_else(aland_km2 > 0, estimate / aland_km2, NA_real_)
  ) %>%
  st_intersection(la_city) %>%  # clip geometries to city boundary for clean edges
  filter(!st_is_empty(geom))

dense_boxes <- get_dense_inset_boxes(la_city, ct_geom, proj_state, proj_county, proj_year, proj_crs, n = 1)
dense1_box  <- dense_boxes$box1
dense1_ct  <- ct_density %>%
  filter(lengths(st_intersects(., dense1_box)) > 0) %>%
  st_intersection(dense1_box)

# Cap colour scale at 97th percentile so outliers don't compress the palette
density_cap <- quantile(ct_density$pop_density, 0.97, na.rm = TRUE)

pop_fill <- list(
  scale_fill_gradientn(
    colours  = c("#fef0d9", "#fdcc8a", "#fc8d59", "#e34a33", "#b30000"),
    limits   = c(0, density_cap),
    oob      = squish,
    na.value = "white",
    name     = "Density\n(pop / km²)",
    labels   = comma_format(accuracy = 1)
  )
)

map1_main <- ggplot() +
  geom_sf(data = ocean,         fill = "#c6e8f5", colour = NA) +
  geom_sf(data = ct_density, aes(fill = pop_density), colour = NA) +
  geom_sf(data = la_water_city, fill = "#c6e8f5", colour = NA) +
  geom_sf(data = dense1_box, fill = NA, colour = "black", linetype="dashed") +
  pop_fill +
  annotation_scale(location = "bl", unit_category = "metric", style = "ticks",
                   line_col = "grey30", text_col = "grey30") +
  coord_sf(
    xlim   = st_bbox(city_view_bbox)[c("xmin", "xmax")],
    ylim   = st_bbox(city_view_bbox)[c("ymin", "ymax")],
    expand = FALSE
  ) +
  labs(title = "Population density") +
  theme_map +
  theme(legend.position="left")

inset_theme <- theme_map + theme(legend.position = "none",
                                 panel.border = element_rect(colour = "black",linewidth = 0.25, linetype = "dashed"))

map1_inset1 <- ggplot() +
  geom_sf(data = dense1_ct, aes(fill = pop_density),
          colour = "white", linewidth = 0.15) +
  geom_sf(data = la_water_city %>% st_intersection(dense1_box),
          fill = "#c6e8f5", colour = NA) +
  pop_fill +
  coord_sf(xlim = st_bbox(dense1_box)[c("xmin", "xmax")],
           ylim = st_bbox(dense1_box)[c("ymin", "ymax")], expand = FALSE) +
  inset_theme



map1 <- ggdraw() +
  draw_plot(map1_main) +
  draw_plot(map1_inset1, x = .7, y = 0.67, width = 0.20, height = 0.28)

ggsave("tables_figures/map1_population_density.png", map1, dpi = 800)
map1


```

## Map 2 — Terrain elevation with hillshade

```{r map-elevation, message=FALSE, warning=FALSE}
# Load DEM (written by download_dem() in 1_data_collection.R)
dem_path <- paste0(base_path, "geo_", proj_county, "/", proj_county, "_dem.tif")
dem_raw  <- terra::rast(dem_path)

# Project to study CRS and crop to city view extent (continuous elevation, no county mask)
dem <- terra::project(dem_raw, paste0("EPSG:", proj_crs))
dem <- terra::crop(dem, terra::vect(city_view_bbox))
names(dem) <- "elevation"

# Hillshade: sun at 40° elevation from NW (direction 300°)
slope  <- terra::terrain(dem, "slope",  unit = "radians")
aspect <- terra::terrain(dem, "aspect", unit = "radians")
hill   <- terra::shade(slope, aspect, angle = 40, direction = 300)
names(hill) <- "hillshade"

map2 <- ggplot() +
  # Hillshade base layer (grayscale); maxcell=Inf prevents resampling artefacts
  geom_spatraster(data = hill, maxcell = Inf) +
  scale_fill_gradientn(
    colours  = grey.colors(256, start = 0, end = 1),
    na.value = NA,
    guide    = "none"
  ) +
  # Elevation colour layer on top of hillshade (semi-transparent for blending)
  new_scale_fill() +
  geom_spatraster(data = dem, alpha = 0.55, maxcell = Inf) +
  scale_fill_hypso_tint_c(
    palette  = "dem_poster",
    na.value = NA,
    name     = "Elevation (m)"
  ) +
  geom_sf(data = st_difference(city_view_bbox, la_city), fill = "white", colour = NA, alpha = 0.5) +
  # Ocean overlaid on top to mask near-zero DEM values in the Pacific
  geom_sf(data = ocean,         fill = "white", colour = NA) +
  geom_sf(data = la_water_city, fill = "white", colour = NA) +
  # City boundary (clipped to county — no ocean overhang)
  annotation_scale(location = "bl", unit_category = "metric", style = "ticks",
                   line_col = "grey30", text_col = "grey30") +
  coord_sf(
    xlim   = st_bbox(city_view_bbox)[c("xmin", "xmax")],
    ylim   = st_bbox(city_view_bbox)[c("ymin", "ymax")],
    expand = FALSE
  ) +
  labs(title = "Terrain elevation") +
  theme_map

ggsave("tables_figures/map2_terrain_elevation.png", map2,
       width = 7, height = 7, dpi = 500)
map2


```

## Combined figure

```{r study-area-combined, fig.width=14, fig.height=7}
combined <- plot_grid(map1, map2, ncol = 2)
ggsave("tables_figures/map_study_area_combined.png", combined,
       width = 14, height = 7, dpi = 0)
combined
```

# Map basic input data

```{r}
source("./0_Libraries.R")
# get county boundary and water
la_county <- get_county_boundary() %>%
  st_transform(proj_crs)

la_city <- get_city_boundary(proj_crs) %>%
  st_transform(proj_crs)

# get water
la_water <- area_water(state=proj_state, county=proj_county)%>%
  st_transform(proj_crs) %>%
  clipintersect_boundary(la_city, buffer = 0)
  
la_food <- foodpoi_plot %>%
  filter(!is.na(LONGITUDE)) %>%
  st_as_sf(coords=c("LONGITUDE", "LATITUDE"), crs=proj_coord_crs) %>%
  st_transform(proj_crs) #%>%
  #clipintersect_boundary(la_city)

la_ct <- get_census_tracts(base_path, proj_crs, state=proj_state, year=proj_year, county=proj_county)

la_ctcent <- get_lac_centroids() %>%
  st_transform(proj_crs) %>% # OSM data is in proj_coord_crs
  mutate(id=row_number()) 

la_ct_wtcent <- get_lac_weight_centroids() %>%
  st_transform(proj_crs) 

osm_parks <- getbb("Los Angeles County", 
                   display_name_contains = "United States") %>%
            opq() %>% 
            add_osm_feature(key = "leisure", 
                  value=c("park", "nature_reserve", "golf_course")) %>% 
            osmdata_sf()

la_parks <- osm_parks$osm_polygons %>%
  st_transform(proj_crs) %>%
  clipintersect_boundary(la_city, buffer = 0)

# get major highways

```

## Mapping study area

```{r}

tm_shape(la_city) +
  tm_polygons(col = "lightgrey", fill_alpha=.2) +
  tm_shape(la_county) +
  tm_borders() + 
  tm_shape(la_water) +
  tm_fill(col="lightblue", lwd=0) +  
  tm_shape(la_parks) +
  tm_polygons(col="lightgreen") + 
  tm_layout(inner.margins = c(0.1, .1, 0.2, 0.1)) +
  tm_scale_bar()

```

## Mapping food access poi

```{r}

tm_shape(la_city) +
  tm_borders(col="darkblue") +
  tm_shape(la_food %>% filter(category!="SPF" | category!="Not included"), alpha=.9) +
  tm_dots(col="category", size = 0.02) +  
  tm_shape(la_water) +
  tm_fill(col="lightblue", lwd=0) + 
  tm_scale_bar(position=c("LEFT", "BOTTOM")) 

```

## Mapping population weighted census tract centroids

```{r}
# remotes::install_version("tmap", version = "3.3-4") fix for tmap view
# detach("package:tmap", unload=TRUE)

tmap_mode("plot")

boundary1 <- st_buffer(la_city, -200) 

la_city_ct <- la_ct %>% filter((lengths(st_intersects(., boundary1)) > 0))

tm_shape(la_city_ct) +
  tm_polygons(col="white") +
  tm_shape(la_ctcent %>% filter((lengths(st_intersects(., la_city)) > 0))) +
  tm_dots(col = "red", size = 0.02) +
  tm_shape(la_ct_wtcent %>% filter((lengths(st_intersects(., la_city)) > 0))) +
  tm_dots(col = "purple", size = 0.02) +  
  tm_shape(la_water) +
  tm_fill(col="lightblue", lwd=0) +
  tm_layout(legend.position = c("left", "bottom")) +
  tm_scale_bar(position = c("left", "bottom"))

```

# Inspecting census tract point datasets for issues

We inspected the census tracts with strange behavior ID = 980016, 990200, 599100 (in the middle of ocean) using census reporter because there were issues plotting these points

980016 - split into two (do not erase water - introduces invalid geoms, handled in code) ✔️

9902 - zero population (remove all cts with 0 aland) ✔️

5991 - catalina island (remove)

```{tract <- la_ct[la_ct$TRACTCE==980016,]}

print(tract$TRACTCE[[1]])  # Find adjacent tracts  

dot <- la_ctcent_dat[la_ctcent_dat$TRACTCE==980016,] 
dotwt <- la_ct_wtcent_dat[la_ct_wtcent_dat$TRACTCE==980016,]

# Plot the selected and adjacent tracts 
figure <- tm_shape(dot) + tm_dots(col = "red", size = 0.01) + 
  tm_shape(dotwt) + 
  tm_dots(col = "green", size = 0.01) + 
  tm_shape(la_ct) + 
  tm_borders() + 
  tm_layout(legend.position = c("left", "bottom"))

print(figure)

```

980030 - no weighted CT point due to zero population (remove all with 0 population) ✔️

we find all empty points in la_ct_wtcent_dat and inspect their corresponding blocks to find they are all population 0; adjusted to remove 0 pop CTs ✔️

```{r}
empty <- la_ct_wtcent_dat[st_is_empty(la_ct_wtcent_dat),]

empty_blocks <- la_cb[la_cb$TRACTCE %in% empty$TRACTCE,]
```

# Mapping POI and population-weighted centroids data

```{r}
tm_shape(la_county) +
  tm_borders() +
  tm_shape(food_23_24_SSI) +
  tm_dots(col = "red", size = 0.01) +
  tm_shape(la_ct) +
  tm_borders() +
  tm_shape(la_ctcent_dat) +
  tm_dots(col = "purple", size = 0.01) +
  tm_layout(legend.position = c("left", "bottom")) +
  tm_scale_bar(position = c("left", "bottom"))

```

## Map terrain elevation and street network features

```{r}

elev <- get_dem("socal", proj_crs)

tmap_options(check.and.fix = TRUE)

tm_shape(elev) +
  tm_raster(palette = terrain.colors(100), style = "cont", title = "Elevation (m)") +
  tm_shape(la_county) +
  tm_borders(col = "black", alpha=.5) +
  tm_scale_bar()


```

## Visualizing OSM

```{r}
  osm_lines <- get_osm(bbox=lac_bbox)
  ht = c("raceway") # highway types to remove
  # highway types https://wiki.openstreetmap.org/wiki/Key:highway#Roads
  osm_paths = osm_lines[!(osm_lines$highway %in% ht), ]
  plot(osm_paths["highway"], key.pos = 1)
  
  rm(osm_lines, osm_paths)
  
  #not enough heap space to run this 
  #network <- street_network_to_sf(r5r_core)


```
