---
title: "3_prelim_results"
author: "Angela Zhang"
date: "2025-04-02"
output:
  pdf_document: default
  html_document: default
---
# Paper Section E: Comparing Food Environment Measures at Different Spatial Scales
# NOTE: This is a standalone RMD file meaning it can run independently of other files in this codebase
#
# Produces comparison statistics across population representation methods (parcel,
# census tract centroid, population-weighted centroid):
#   - CV (coefficient of variation): within-tract variation scaled by mean
#   - Mean difference (bias): systematic directional error
#   - MAE: mean absolute error (typical absolute disagreement)
#   - RMSE: root mean squared error (weights larger errors more)
#   - Paired t-tests and Wilcoxon rank-sum tests for differences in distributions


```{r}
source("0_Libraries.R")
library(kableExtra)
library(tidycensus)
library(Metrics)
source("./helper/data_functions.R")
library(cowplot)

la_city <- get_city_boundary(proj_crs) 

la_hh <- get_lac_households(processed_path, proj_coord_crs)
 
la_ct <- get_census_tracts(base_path, proj_crs, state=proj_state, year=proj_year, county=proj_county) |>
  mutate(GEOID=as.character(GEOID)) %>%
  st_as_sf()

# get census tracts intersecting with LA city boundary 
# is this how the la_ct is typically clipped? 
la_city_ct <- la_ct %>%
  st_transform(proj_crs) %>%
  clipintersect_boundary(la_city, buffer = -50)  

# include households with census tract in la city
la_city_hh <- la_hh %>%
  filter(GEOID_20 %in% la_city_ct$GEOID)

# get water
la_water <- area_water(state=proj_state, county=proj_county) %>%
  st_transform(proj_crs) 

la_openspace <- get_openspace_parks() |> 
    st_transform(proj_crs) 


# osm_parks <- getbb("Los Angeles County", display_name_contains = "United States") %>%
#   opq() %>% 
#   add_osm_feature(key = "leisure", value=c("park", "nature_reserve", "golf_course")) %>%
#   osmdata_sf()
# 
# la_parks <- osm_parks$osm_polygons %>%
#   st_transform(proj_crs) %>%
#   clipintersect_boundary(la_city, buffer = 0)


```
# Sanity checks
✔️ We expect there to be 1424424 households in LA City
✔️ We also want to see why some CTs are missing observations in the household aggregate measures
  - No households were located in those CTs 
  - Want to check why that is (is the population zero? did they get misattributed to the wrong CTs?)

```{r}
nrow(la_city_hh) # 1424424 households 
length(unique(la_city_hh$GEOID_20))
nrow(la_city_ct) # 1303 census tracts

sum(la_city_hh$Units_12345_2023)
```

### Load CT level data for all measures

```{r}

# TODO un-transform GEOID from as numeric 
# visualize ct_level data
# head(read.csv(paste0(processed_path, "/LAC_cleaned/ct_driving_times.csv"))) 
#'* Data frame with household values of food environment measures merged with census data*
parcel_data0 <- fread(paste0(processed_path, "LAC_cleaned/parcel_driving_all.csv"))

parcel_data_all <- parcel_data0 |> 
  tidytable::left_join(la_city_hh, by.x = c("id", "GEOID"), by.y = c("id", "GEOID_20"))


#'* Data frame with all food environment measures aggregated to the census tract level (geographic specificity) *
ct_drive_data <- la_ct %>%
  merge(read_csv(paste0(processed_path, "/LAC_cleaned/ct_driving_times.csv")), by= "GEOID")



# temp <- read_csv(paste0(processed_path, "LAC_cleaned/dt_household_ct.csv"))

```

### (optional) Robustness check: Inspecting missing census tract data in the household dataset

Missing summary: 
- 11 tracts in la_city_ct but not in la_city_hh; tracts are large and often near edges 
- GEOIDs are:  06037980031 06037980033 06037980030 06037980039 06037980017 06037980037 6037980002 6037980005 6037980020 6037265301
- defining la_city_hh using spatial boundary rather than filtering by GEOID does NOT make a difference (it's exactly the same)
```{r, eval=FALSE}
missing <- setdiff(as.numeric(la_city_ct$GEOID), as.numeric(ct_drive_data$GEOID))
missing2 <- setdiff(as.numeric(la_city_ct$GEOID), as.numeric(la_city_hh$GEOID_20))

# TODo map census tracts with all NA values for driving_SMK_15_parcel_cv_q
res <- la_city_ct %>% filter(as.numeric(GEOID) %in% missing2)
print(missing2)

# plot census tracts that are missing from parcel dataset
tm_shape(la_city) +
  tm_polygons(col="yellow") + 
  tm_shape(la_city_ct[as.numeric(la_city_ct$GEOID) %in% missing2,] ) + 
  tm_polygons(col= "red", alpha=.5) 

# try defining la_city_hh using spatial boundary rather than filtering tracts
temp <- la_hh |> 
  st_transform(proj_crs) |>
  clipintersect_boundary(la_city, buffer = 0)

# we find that it's the same


#'*Why NA values for ct_wtcent of a particular census tract * 
missing_wtct <- parcel_data_all %>% filter(is.na(ct_wtcent))
original_wtct <- dt_ct_wtcentm %>% filter(GEOID %in% missing_wtct$GEOID) 
# we find that missing tract is population = 0 https://datacommons.org/place/geoId/06037980022
# should be CT: 106645 (misgeocoded)
data_to_conv <- parcel_data_all %>% filter(GEOID==6037980022)

```

```{r robustness-na-cv-map, message=FALSE, warning=FALSE}
# CTs where CV for SMK 15-min is NA (zero mean → CV undefined)
na_cv_cts <- parcel_data_ct %>%
  filter(is.na(driving_SMK_15_parcel_cv))

stopifnot(nrow(na_cv_cts)==0)

# this map should be empty
ggplot() +
  geom_sf(data = la_ct[la_ct$GEOID %in% na_cv_cts$GEOID,], fill = "red", colour = "red") +
  geom_sf(data = la_city, color = "grey")
  labs(title = paste0("CTs with NA CV — SMK 15-min drive (n=", nrow(na_cv_cts), ")")) +
  theme_map
```

# Remove large variables that are no longer needed for analysis
```{r}
rm(la_city_hh)
rm(la_hh)
```


# Data analysis

## Calculate metrics and agreements btwn parcel-level and census tract-level values
- Pearson's R (linear correlation)
- Spearman Rank p (non parametric correlation between variables)
- Bias
- MAE
- RMSE 
```{r}
# sub <- parcel_data_all |> 
#   filter(type=="SMK" & drive==15)

results <- parcel_data_all |>
  group_by(type, drive, network) |> 
  summarize(bias_hh_ct = bias(count, ct_cent), 
            mae_hh_ct = mae(count, ct_cent), 
            rmse_hh_ct = rmse(count, ct_cent),
            bias_hh_wct = bias(count, ct_wtcent),
            mae_hh_wct = mae(count, ct_wtcent), 
            rmse_hh_wct = rmse(count, ct_wtcent),
            bias_wct_ct = bias(ct_wtcent, ct_cent),
            mae_wct_ct = mae(ct_wtcent, ct_cent), 
            rmse_wct_ct = rmse(ct_wtcent, ct_cent))

results

data.table::fwrite(results, "./tables_figures/figure4_data.csv")


```
## Plot histograms to observe distributions 
We find that many variables are not normally distributed so that leads us to select the Wilcoxon rank sum test to test differences in means between the various groups

*Todo* Additionally, we want to try to identify what type of measurement error/bias this could lead to which requires observing the distributions of the parcel-level values compared to the aggregate values (among other techniques). The figures below are a start

*Todo* look into the zero values for CT_CENT that are shown in the third figure
```{r}
dist <- ct_drive_data |>
  as.data.table() |>
  melt(measure.vars=c("ct_cent", "ct_wtcent", "parcel_mean")) 

fig_hist_all <- ggplot(dist, aes(x=value, fill=variable, alpha=.5)) +
   geom_step(stat = "bin", bins = 50, direction = "mid") +
  geom_histogram(bins = 50, alpha = 0.3, colour = NA,
    position = "identity") +
  facet_grid(cols=vars(type), rows=vars(drive), scales="free") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
ggsave("./tables_figures/fig_hist_all_measures.png", fig_hist_all, width = 14, height = 10, dpi = 300)
fig_hist_all

fig_hist_smk <- ggplot(dist |> filter(type=="SMK"), aes(x=value, col=variable, fill=variable, alpha=.3)) +
  geom_step(stat = "bin", bins = 50, direction = "mid") +
  geom_histogram(bins = 50, alpha = 0.3, colour = NA,
    position = "identity") +
  #geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  facet_grid(rows=vars(drive), scales="free") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
ggsave("./tables_figures/fig_hist_smk.png", fig_hist_smk, width = 7, height = 10, dpi = 300)
fig_hist_smk

```

## Distribution of parcel, centroid, and weighted centroid measures (faceted by drive time × retailer)
```{r dist-all-measures, fig.width=14, fig.height=10}
measure_labels <- c(
  parcel_mean = "Parcel (ground truth)",
  ct_cent     = "CT centroid",
  ct_wtcent   = "Pop.-weighted centroid"
)

dist_long <- ct_drive_data |>
  st_drop_geometry() |>
  filter(type %in% c("CNV", "FF", "GRC", "RR", "SMK", "SPF")) |>
  pivot_longer(c(parcel_mean, ct_cent, ct_wtcent),
               names_to = "measure", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(
    measure = factor(measure,
                     levels = c("parcel_mean", "ct_cent", "ct_wtcent"),
                     labels = unname(measure_labels)),
    drive_label = paste0(drive, " min")
  )

fig_dist_all <- ggplot(dist_long, aes(x = value, colour = measure, fill = measure)) +
  geom_density(alpha = 0.20, linewidth = 0.6) +
  facet_wrap(~type+drive_label,
    scales = "free", 
    ncol = 6
  ) +
  scale_colour_manual(values = c("Parcel (ground truth)" = "#2166ac",
                                 "CT centroid"           = "#d6604d",
                                 "Pop.-weighted centroid"= "#4dac26")) +
  scale_fill_manual(  values = c("Parcel (ground truth)" = "#2166ac",
                                 "CT centroid"           = "#d6604d",
                                 "Pop.-weighted centroid"= "#4dac26")) +
  labs(
    x      = "Accessibility (count within threshold)",
    y      = "Density",
    colour = "Measure",
    fill   = "Measure",
    title  = "Distribution of accessibility measures by drive time and retailer type") +
  theme_bw(base_size = 15) +
  theme(
    legend.position   = "bottom",
    strip.text        = element_text(size = 10, face = "bold"),
    axis.text.x       = element_text(size = 10, angle = 45, vjust = 1, hjust = 1),
    panel.grid.minor  = element_blank()
  )
ggsave("./tables_figures/fig_dist_all_measures.png", fig_dist_all, width = 14, height = 10, dpi = 500)
fig_dist_all
```
## Calculate t-tests to evaluate differences in distributions for census tract metrics and parcel metrics aggregated to census level 
```{r}
i=1
ttests0 <- split(ct_drive_data, list(ct_drive_data$type, ct_drive_data$drive)) %>%
  lapply(
  function(df) { 
    i <<- i+1
    print(length(df$parcel_mean))
      # paired n for each comparison (drops NAs pairwise)
    n1 <- sum(complete.cases(df$parcel_mean, df$ct_cent))
    n2 <- sum(complete.cases(df$ct_wtcent, df$ct_cent))
    n3 <- sum(complete.cases(df$parcel_mean, df$ct_wtcent))
    
    t1 <<- t.test(df$parcel_mean, df$ct_cent, paired=TRUE, alternative="two.sided")
    t2 <- t.test(df$ct_wtcent, df$ct_cent, paired=TRUE, alternative="two.sided")
    t3 <- t.test(df$parcel_mean, df$ct_wtcent, paired =TRUE, alternative="two.sided")

    
    drive <- unique(df$drive)[1]
    print(drive)
    type <- unique(df$type)[1]
    print(type)
    res <- data.frame(
      time_dist = unique(df$drive)[1], 
      type =  unique(df$type)[1],
      tests = c("parcel_cent", "wtcent_cent", "parcel_wtcent"), 
      n = c(n1, n2, n3),
      pval = c(t1$p.value, t2$p.value, t3$p.value),
      est = c(t1$estimate, t2$estimate, t3$estimate),
         conf.low  = c(t1$conf.int[1], t2$conf.int[1], t3$conf.int[1]),
      conf.high = c(t1$conf.int[2], t2$conf.int[2], t3$conf.int[2])
    )
    print(length(res))
    return(res)
  }
)

ttests <- ttests0 |>
  bind_rows()

ttests %>%
  select(type, time_dist, tests, n, est, conf.low, conf.high, pval) %>%
  kable(digits = 4, caption = "Paired t-test results") %>%
    row_spec(0, bold = TRUE, background = "white", color = "black") 
```
## Plotting results of t-test 
The results of the t-test shows that for longer drive time distances (>25 minutes), means of the census-tract level parcel measure are significantly different from means of the centroid calculated measure and means of population-centroid calculated measures 
```{r}
fig_ttest <- ggplot(ttests, aes(x=fct_relevel(tests, "parcel_cent", "wtcent_cent", "parcel_wtcent"), y=est, color=(pval<.05))) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  facet_grid(cols=vars(time_dist), rows=vars(type), scales="free") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

ggsave("./tables_figures/fig_ttest_results.png", fig_ttest, width = 14, height = 10, dpi = 300)
fig_ttest

fig_scatter_parcel_wtcent <- ggplot(ct_drive_data, aes(x=parcel_mean, y=ct_wtcent)) +
  geom_point(size=1, aes(alpha=.5)) +
  facet_grid(cols=vars(drive), rows=vars(type), scales="fixed") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
ggsave("./tables_figures/fig_scatter_parcel_wtcent.png", fig_scatter_parcel_wtcent, width = 14, height = 10, dpi = 300)
fig_scatter_parcel_wtcent

unique(ct_drive_data$type)

# TODO make table of the t-tests
# TODO make figure of t-tests

```
## Wilcox test results
Given that distributions of counts are not strictly following a normal distribution, we also perform a Wilcox test
```{r}
wtests0 <- split(ct_drive_data, list(ct_drive_data$type, ct_drive_data$drive)) %>%
  lapply(
  function(df) { 
   # i <<- i+1
    print(length(df$parcel_mean))
      # paired n for each comparison (drops NAs pairwise)
    n1 <- sum(complete.cases(df$parcel_mean, df$ct_cent))
    n2 <- sum(complete.cases(df$ct_wtcent, df$ct_cent))
    n3 <- sum(complete.cases(df$parcel_mean, df$ct_wtcent))

    
    t1 <<- wilcox.test(df$parcel_mean, df$ct_cent, paired=TRUE, alternative="two.sided")
    t2 <- wilcox.test(df$ct_wtcent, df$ct_cent, paired=TRUE, alternative="two.sided")
    t3 <- wilcox.test(df$parcel_mean, df$ct_wtcent, paired =TRUE, alternative="two.sided")

    drive <- unique(df$drive)[1]
    print(drive)
    type <- unique(df$type)[1]
    print(type)
    res <- data.frame(
      time_dist = unique(df$drive)[1], 
      type =  unique(df$type)[1],
      tests = c("parcel_cent", "wtcent_cent", "parcel_wtcent"), 
      n = c(n1, n2, n3),
      pval = c(t1$p.value, t2$p.value, t3$p.value)
      #est = c(t1$estimate, t2$estimate, t3$estimate),
      #conf.low  = c(t1$conf.int[1], t2$conf.int[1], t3$conf.int[1]),
      #conf.high = c(t1$conf.int[2], t2$conf.int[2], t3$conf.int[2])
    )
    print(length(res))
    return(res)
  }
)

wtests <- wtests0 |>
  bind_rows() |> 
  pivot_wider(values_from=pval, names_from=c(tests,n)) %>%
  select(type, time_dist, everything())%>%
  arrange(type)

wtests %>%
  #select(type, time_dist, tests, n, pval) %>%
  kable(digits = 4, caption = "Paired Wilcox (Mann Whitney U)-test results", booktabs = TRUE) %>%
  row_spec(0, bold = TRUE, background = "white", color = "black") |>
  kableExtra::collapse_rows(valign = "top")

unique(ct_drive_data$type)
gc()

```

# Mapping data 
## Map variation within CTs using parcel data
- map parcel data for one rural tract, one urban tract, one suburban tract
*todo* figure out why there are missing values for 5 minutes
- map coefficient of variation (dispersion) for each CT in LA City using parcel data
### Setup census tract parcel data 
```{r}
#'* Data frame with parcel means of food environment measures only: used for demonstrating variation within census tracts *
parcel_data_ct <- read_csv(paste0(processed_path, "/LAC_cleaned/dt_household_ct.csv")) |>
  filter(GEOID %in% la_city_ct$GEOID) |>
  merge(la_city_ct, by= "GEOID") |>
  st_as_sf() |>
  st_transform(proj_crs)

# Derive CV = SD / mean * 100
sd_cols <- grep("_parcel_sd$", names(parcel_data_ct), value = TRUE)
for (col in sd_cols) {
  print(col)
  parcel_data_ct[[sub("_sd$", "_cv", col)]] <-
    parcel_data_ct[[col]] / parcel_data_ct[[sub("_sd$", "_mean", col)]] * 100
}

class(parcel_data_ct)

```

### Setup mapping functions 

```{r cv-maps, message=FALSE, warning=FALSE}
library(scales)
library(ggspatial)
 
# Compute CV quartiles for all type × drive combinations
# (used in maps below and Table 4 — moved earlier from make_table4)
cv_cols <- grep("^driving_.+_\\d+_parcel_cv$", names(parcel_data_ct), value = TRUE)

for (cv_col in cv_cols) {
  q_col  <- sub("_cv$", "_cv_q", cv_col)
  vals   <- parcel_data_ct[[cv_col]]
  breaks <- quantile(vals, c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
  labels <- c(
    sprintf("Q1 (%.0f–%.0f)", breaks[1], breaks[2]),
    sprintf("Q2 (%.0f–%.0f)", breaks[2], breaks[3]),
    sprintf("Q3 (%.0f–%.0f)", breaks[3], breaks[4]),
    sprintf("Q4 (%.0f–%.0f)", breaks[4], breaks[5])
  )
  parcel_data_ct[[q_col]] <- factor(ntile(vals, 4), labels = labels)
}

# Map view extent and ocean (matching 1_visualizations.Rmd style)
city_view_bbox <- st_as_sfc(st_bbox(st_buffer(la_city, 10000))) %>% st_set_crs(proj_crs)

ca    <- tigris::states(cb = TRUE, class = "sf", progress_bar = FALSE) %>%
            dplyr::filter(STUSPS == "CA") %>%
            st_transform(proj_crs)

ocean     <- st_difference(city_view_bbox, st_union(ca))
dense_boxes <- get_dense_inset_boxes(la_city, la_ct, proj_state, proj_county, proj_year, proj_crs, buffer_sizes = c(6000, 4000, 4500))
 
dense_box1  <- dense_boxes$box1
dense_box2  <- dense_boxes$box2
dense_box3  <- dense_boxes$box3

# Clip tracts and water to city boundary and to each inset box
ct_clipped   <- parcel_data_ct  %>%
  clipintersect_boundary(la_city, buffer = 0)

ct_inset1    <- ct_clipped %>%
  clipintersect_boundary(dense_box1, buffer = 0)

ct_inset2    <- ct_clipped %>%
  clipintersect_boundary(dense_box2, buffer = 0)

ct_inset3    <- ct_clipped %>%
  clipintersect_boundary(dense_boxes$box3, buffer = 0)

water_inset1 <- la_water  %>%
  clipintersect_boundary(dense_box1, buffer = 0)

water_inset2 <- la_water  %>%
  clipintersect_boundary(dense_box2, buffer = 0)

water_inset3 <- la_water  %>%
  clipintersect_boundary(dense_box3, buffer = 0)

park_inset1 <- la_openspace %>%
  clipintersect_boundary(dense_box1, buffer = 0)

park_inset2 <- la_openspace %>%
  clipintersect_boundary(dense_box2, buffer = 0)

park_inset3 <- la_openspace %>%
  clipintersect_boundary(dense_box3, buffer = 0)

make_cv_quartile_fill <- function(q_col) {
  lvls <- levels(parcel_data_ct[[q_col]])
  scale_fill_manual(
    values = setNames(c("#e9d8a6", "#ee9b00", "#ca6702", "#ae2012"), lvls),
    name   = "CV Quartile (%)",
    drop   = FALSE
  )
}

inset_theme <- theme_map + theme(
  legend.position = "none",
  panel.border    = element_rect(colour = "black", linewidth = 0.25, linetype = "dashed")
)

make_cv_map <- function(col, m_col, title, fill_scale) {
  main <- ggplot() +
    #geom_sf(data = la_city, fill = "#f4f1de") +
    geom_sf(data = ct_clipped,  aes(fill = .data[[col]]), colour = NA) +
      fill_scale

  zero_means <- ct_clipped %>% filter(.data[[m_col]]==0)

  if(nrow(zero_means) > 0) {
    main <- main + geom_sf(data = subset(ct_clipped, ct_clipped[[m_col]]==0), fill="grey", colour = NA, name="Weighted mean equals zero")
  }

  main <- main +
    geom_sf(data = la_water,    fill = "#005f73",   colour = NA) +
    geom_sf(data = la_openspace, aes(colour="Parks") ,fill="#94d2bd", alpha = 0.5) +
    scale_color_manual(values=c("Parks" = "#94d2bd")) +
    geom_sf(data = dense_box1,  fill = NA, colour = "black", linetype = "dashed", linewidth = 0.4) +
    geom_sf(data = ocean,       fill = "#005f73", colour = NA) +
    geom_sf(data = dense_box2,  fill = NA, colour = "black", linetype = "dashed", linewidth = 0.4) +
    geom_sf(data = dense_box3,  fill = NA, colour = "black", linetype = "dashed", linewidth = 0.4) +
    annotation_scale(location = "bl", unit_category = "metric", style = "ticks",
                     line_col = "grey30", text_col = "grey30") +
    coord_sf(xlim = st_bbox(city_view_bbox)[c("xmin", "xmax")],
             ylim = st_bbox(city_view_bbox)[c("ymin", "ymax")], expand = FALSE) +
    labs(title = title) +
    theme_map + theme(legend.position = "left", panel.background = element_rect(fill = "#f4f1de"))

  make_inset <- function(ct_data, water_data, park_data, box) {
    inset <- ggplot() +
      geom_sf(data = ct_data,    aes(fill = .data[[col]]), colour = "white", linewidth = 0.15)

    if(nrow(zero_means) > 0) {
      inset <- inset + geom_sf(data = subset(ct_clipped, ct_clipped[[m_col]]==0), fill="grey", colour = NA, name="Weighted mean equals zero")
    }
    inset <- inset +
      geom_sf(data = water_data, fill = "#005f73", colour = NA) +
      geom_sf(data = park_data, fill = "#94d2bd", colour = NA, alpha = 0.5) +
      fill_scale +
      coord_sf(xlim = st_bbox(box)[c("xmin", "xmax")],
               ylim = st_bbox(box)[c("ymin", "ymax")], expand = FALSE) +
      inset_theme
  }

  ggdraw() +
    draw_plot(main) +
    draw_plot(make_inset(ct_inset1, water_inset1, park_inset1, dense_box1), x = 0.74, y = 0.67, width = 0.28, height = 0.32) +
    draw_plot(make_inset(ct_inset2, water_inset2, park_inset2, dense_box2), x = 0.74, y = 0.37, width = 0.20, height = 0.25) +
    draw_plot(make_inset(ct_inset3, water_inset3, park_inset3, dense_box3), x = 0.74, y = 0.07, width = 0.20, height = 0.25)
}
```

### Make CV maps by quartile
```{r}
map_cv_smk5  <- make_cv_map("driving_SMK_5_parcel_cv_q",  "driving_SMK_5_parcel_mean",  "Supermarket access (5-min)",  fill_scale = make_cv_quartile_fill("driving_SMK_5_parcel_cv_q"))
map_cv_smk10 <- make_cv_map("driving_SMK_10_parcel_cv_q", "driving_SMK_10_parcel_mean", "Supermarket access (10-min)", fill_scale = make_cv_quartile_fill("driving_SMK_10_parcel_cv_q"))
map_cv_smk15 <- make_cv_map("driving_SMK_15_parcel_cv_q", "driving_SMK_15_parcel_mean", "Supermarket access (15-min)", fill_scale = make_cv_quartile_fill("driving_SMK_15_parcel_cv_q"))

map_cv_smk5
map_cv_smk10
map_cv_smk15

ggsave("./tables_figures/fig_cv_smk_5min.png",  map_cv_smk5,  dpi = 500, width = 12, bg = "white")
ggsave("./tables_figures/fig_cv_smk_10min.png", map_cv_smk10, dpi = 500, width = 12, bg = "white")
ggsave("./tables_figures/fig_cv_smk_15min.png", map_cv_smk15, dpi = 500, width = 12, bg = "white")

gc()

```

## CV distribution across LA City census tracts
Distribution of CV (%) by retailer type and drive-time cutoff, across all CTs with non-missing values.

```{r cv-distribution-table}
cv_dist_tbl <- lapply(cv_cols, function(cv_col) {
  vals  <- parcel_data_ct[[cv_col]]
  vals  <- vals[!is.na(vals)]
  type  <- sub("^driving_(.+)_(\\d+)_parcel_cv$", "\\1", cv_col)
  drive <- as.integer(sub("^driving_(.+)_(\\d+)_parcel_cv$", "\\2", cv_col))
  data.frame(
    Type           = type,
    `Drive (min)`  = drive,
    n              = length(vals),
    `Median [IQR]` = sprintf("%.1f [%.1f–%.1f]", median(vals), quantile(vals, 0.25), quantile(vals, 0.75)),
    Range          = sprintf("%.1f–%.1f", min(vals), max(vals)),
    check.names    = FALSE
  )
}) %>%
  bind_rows() %>%
  arrange(Type, `Drive (min)`)

kable(cv_dist_tbl,
      caption = "Distribution of CV (%) across LA City census tracts, by retailer type and drive-time cutoff",
      align   = c("l", "r", "r", "r", "r")) %>%
  kable_styling(full_width = FALSE)
```

## Table 4: Census-tract features by quartiles of coefficient of variation
Pull ACS demographic/socioeconomic features and summarize by CV quartile. CV is based on the within-CT dispersion of parcel-level supermarket accessibility (15-min drive).

### Load ACS
```{r table4-cv-quartiles}
library(tidycensus)
# TODO missing CVs in some CTs (show up as NA) 06037300301

# ACS variables: population, income, poverty, education, Hispanic share
acs_vars <- c(
  "B01003_001", # Total population
  "B19013_001", # Median household income
  "B17001_001", # Poverty universe
  "B17001_002", # Below poverty level
  "B15003_001", # Population 25+ (education denominator)
  "B15003_022", # Bachelor's degree
  "B15003_023", # Master's degree
  "B15003_024", # Professional school degree
  "B15003_025", # Doctorate degree
  "B03003_001", # Total population (Hispanic origin table)
  "B03003_003"  # Hispanic or Latino
)

ct_acs <- get_acs(
  geography = "tract",
  variables = acs_vars,
  state = proj_state,
  year = proj_year,
  output = "wide",
  geometry = FALSE
) %>%
  transmute(
    GEOID        = GEOID,
    pop_total    = B01003_001E,
    hhinc_median = B19013_001E,
    pct_poverty  = B17001_002E / B17001_001E * 100,
    pct_college  = (B15003_022E + B15003_023E + B15003_024E + B15003_025E) / B15003_001E * 100,
    pct_hispanic = B03003_003E / B03003_001E * 100
  )

# Land/water area from tigris (ALAND, AWATER columns present in la_ct)
ct_area <- la_ct %>%
  st_drop_geometry() %>%
  transmute(
    GEOID      = as.character(GEOID),
    aland_km2  = ALAND / 1e6,
    awater_km2 = AWATER / 1e6
  )
```
### Plot tables
```{r}

# Formatting helpers (defined once, used inside make_table4)
fmt_meansd  <- function(x) sprintf("%.1f ± %.1f", mean(x, na.rm = TRUE), sd(x, na.rm = TRUE))
fmt_med_iqr <- function(x) sprintf("%.0f [%.0f–%.0f]",
  median(x, na.rm = TRUE), quantile(x, .25, na.rm = TRUE), quantile(x, .75, na.rm = TRUE))
fmt_pval    <- function(p) if (p < 0.001) "<0.001" else sprintf("%.3f", p)

# Build and display Table 4 for one retailer type × drive-time combination
make_table4 <- function(type, drive, parcel_data_ct, ct_acs, ct_area) {
  cv_col <- paste0("driving_", type, "_", drive, "_parcel_cv")
  q_col  <- paste0("driving_", type, "_", drive, "_parcel_cv_q")
  if (!cv_col %in% names(parcel_data_ct)) return(NULL)

  ct_cv_tbl <- parcel_data_ct %>%
    st_drop_geometry() %>%
    transmute(
      GEOID,
      cv_value    = .data[[cv_col]],
      cv_quartile = .data[[q_col]]
    ) %>%
    left_join(ct_acs, by = "GEOID") %>%
    left_join(ct_area, by = "GEOID") %>%
    mutate(pop_density = pop_total / aland_km2)

  table4 <- ct_cv_tbl %>%
    group_by(cv_quartile) %>%
    summarise(
      `N tracts`                             = n(),
      `CV, mean ± SD`                        = fmt_meansd(cv_value),
      `Population, median [IQR]`             = fmt_med_iqr(pop_total),
      `Pop. density (per km²), median [IQR]` = fmt_med_iqr(pop_density),
      `Land area (km²), median [IQR]`        = sprintf("%.2f [%.2f–%.2f]",
          median(aland_km2, na.rm = TRUE),
          quantile(aland_km2, .25, na.rm = TRUE),
          quantile(aland_km2, .75, na.rm = TRUE)),
      `% Hispanic, mean ± SD`                = fmt_meansd(pct_hispanic),
      `% Below poverty, mean ± SD`           = fmt_meansd(pct_poverty),
      `% College educated, mean ± SD`        = fmt_meansd(pct_college),
      `Median HH income, median [IQR]`       = sprintf("$%.0f [$%.0f–$%.0f]",
          median(hhinc_median, na.rm = TRUE),
          quantile(hhinc_median, .25, na.rm = TRUE),
          quantile(hhinc_median, .75, na.rm = TRUE))
    ) %>%
    rename(`CV Quartile` = cv_quartile)

  kw_pvals <- tibble::tribble(
    ~Variable,                                ~`p-value`,
    "N tracts",                               NA_character_,
    "CV, mean ± SD",                          fmt_pval(kruskal.test(cv_value     ~ cv_quartile, ct_cv_tbl)$p.value),
    "Population, median [IQR]",               fmt_pval(kruskal.test(pop_total    ~ cv_quartile, ct_cv_tbl)$p.value),
    "Pop. density (per km²), median [IQR]",   fmt_pval(kruskal.test(pop_density  ~ cv_quartile, ct_cv_tbl)$p.value),
    "Land area (km²), median [IQR]",          fmt_pval(kruskal.test(aland_km2    ~ cv_quartile, ct_cv_tbl)$p.value),
    "% Hispanic, mean ± SD",                  fmt_pval(kruskal.test(pct_hispanic ~ cv_quartile, ct_cv_tbl)$p.value),
    "% Below poverty, mean ± SD",             fmt_pval(kruskal.test(pct_poverty  ~ cv_quartile, ct_cv_tbl)$p.value),
    "% College educated, mean ± SD",          fmt_pval(kruskal.test(pct_college  ~ cv_quartile, ct_cv_tbl)$p.value),
    "Median HH income, median [IQR]",         fmt_pval(kruskal.test(hhinc_median ~ cv_quartile, ct_cv_tbl)$p.value)
  )

  table4 %>%
    mutate(`N tracts` = as.character(`N tracts`)) %>%
    pivot_longer(-`CV Quartile`, names_to = "Variable", values_to = "val") %>%
    pivot_wider(names_from = `CV Quartile`, values_from = val) %>%
    left_join(kw_pvals, by = "Variable") %>%
    select(Variable, everything())
}

# Detect all available type × drive combinations from parcel_data_ct column names
cv_cols <- grep("^driving_.+_\\d+_parcel_cv$", names(parcel_data_ct), value = TRUE)
combos  <- data.frame(
  type  = sub("^driving_(.+)_(\\d+)_parcel_cv$", "\\1", cv_cols),
  drive = as.integer(sub("^driving_(.+)_(\\d+)_parcel_cv$", "\\2", cv_cols))
) %>% arrange(type, drive)

# Generate, display, and save Table 4 for every combination
table4_list <- list()
for (i in seq_len(nrow(combos))) {
  type  <- combos$type[i]
  drive <- combos$drive[i]
  tbl   <- make_table4(type, drive, parcel_data_ct, ct_acs, ct_area)
  if (is.null(tbl)) next

  caption_str <- sprintf(
    "Table 4. CT features by CV quartile (%s, %d-min drive). p from Kruskal-Wallis.",
    type, drive
  )
  print(
    tbl %>%
      kable(caption = caption_str,
            align   = c("l", rep("r", ncol(tbl) - 1))) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
      row_spec(0, bold = TRUE)
  )

  data.table::fwrite(tbl, sprintf("./tables_figures/table4_%s_%d.csv", type, drive))
  table4_list[[i]] <- tbl %>% mutate(type = type, drive = drive, .before = 1)
}

table4_combined <- bind_rows(table4_list)
data.table::fwrite(table4_combined, "./tables_figures/table4_combined.csv")
```

## Bland-Altman Analysis: Centroid vs. Parcel Agreement

```{r bland-altman, fig.width=12, fig.height=10}
library(blandr)

# Long-form: one row per CT × retailer × drive × centroid type
ba_data <- ct_drive_data %>%
  st_drop_geometry() %>%
  filter(!is.na(parcel_mean)) %>%
  pivot_longer(c(ct_cent, ct_wtcent), names_to = "centroid_type", values_to = "centroid") %>%
  filter(!is.na(centroid)) %>%
  mutate(
    diff = centroid - parcel_mean,
    mean = (centroid + parcel_mean) / 2,
    centroid_label = factor(centroid_type,
      levels = c("ct_cent", "ct_wtcent"),
      labels = c("Unweighted centroid", "Pop.-weighted centroid"))
  )

# BA statistics + proportional bias per group (one blandr call per group)
# CIs follow Bland & Altman (1986/1999): SE(bias) = s/sqrt(n), SE(LoA) = sqrt(3)*s/sqrt(n),
# using t-distribution with n-1 df — as implemented in blandr.statistics()
ba_summary <- ba_data %>%
  group_by(centroid_label, type, drive) %>%
  group_modify(~ {
    s   <- blandr.statistics(.x$centroid, .x$parcel_mean)
    fit <- lm(diff ~ mean, data = .x)
    tibble(
      n          = nrow(.x),
      bias       = s$bias,       bias_lo  = s$biasLowerCI,      bias_hi  = s$biasUpperCI,
      loa_lo     = s$lowerLOA,   loa_lo_lo = s$lowerLOA_lowerCI, loa_lo_hi = s$lowerLOA_upperCI,
      loa_hi     = s$upperLOA,   loa_hi_lo = s$upperLOA_lowerCI, loa_hi_hi = s$upperLOA_upperCI,
      slope      = coef(fit)["mean"],
      slope_p    = coef(summary(fit))["mean", "Pr(>|t|)"]
    )
  }) %>%
  mutate(slope_label = sprintf("\u03b2=%.3f%s", slope, if_else(slope_p < 0.05, "*", "")))

# Helper: BA plot for one centroid type, faceted by retailer × drive time
# CI bands on bias and LoA lines per Bland & Altman (1999)
ba_plot <- function(centroid_lbl) {
  d    <- filter(ba_data,    centroid_label == centroid_lbl)
  smry <- filter(ba_summary, centroid_label == centroid_lbl)

  ggplot(d, aes(mean, diff)) +
    geom_point(alpha = 0.2, size = 0.6, colour = "grey40") +
    # 95% CI bands for bias and LoA (shaded rectangles spanning full x range)
    geom_rect(data = smry, aes(xmin=-Inf, xmax=Inf, ymin=bias_lo,  ymax=bias_hi),
              fill = "steelblue", alpha = 0.15, inherit.aes = FALSE) +
    geom_rect(data = smry, aes(xmin=-Inf, xmax=Inf, ymin=loa_lo_lo, ymax=loa_lo_hi),
              fill = "tomato",    alpha = 0.15, inherit.aes = FALSE) +
    geom_rect(data = smry, aes(xmin=-Inf, xmax=Inf, ymin=loa_hi_lo, ymax=loa_hi_hi),
              fill = "tomato",    alpha = 0.15, inherit.aes = FALSE) +
    # Bias and LoA lines
    geom_hline(data = smry, aes(yintercept = bias),   colour = "steelblue", linewidth = 0.8) +
    geom_hline(data = smry, aes(yintercept = loa_lo), colour = "tomato", linetype = "dashed") +
    geom_hline(data = smry, aes(yintercept = loa_hi), colour = "tomato", linetype = "dashed") +
    # Proportional bias regression line
    geom_smooth(method = "lm", se = TRUE, colour = "darkorange", linewidth = 0.6, alpha = 0.15) +
    geom_text(data = smry, aes(x = -Inf, y = Inf, label = slope_label),
              hjust = -0.1, vjust = 1.5, size = 2.5, colour = "darkorange4") +
    geom_hline(yintercept = 0, linewidth = 0.3) +
    facet_grid(type ~ drive, labeller = labeller(drive = \(x) paste0(x, " min")), scales = "free") +
    labs(x = "Mean of centroid and parcel", y = "Centroid \u2212 Parcel",
         title = paste("Bland-Altman:", centroid_lbl, "vs. parcel (ground truth)"),
         subtitle = "Blue = bias \u00b1 95% CI  |  Red dashed = 95% LoA \u00b1 95% CI  |  Orange = proportional-bias fit (* p<0.05)") +
    theme_bw(base_size = 9) +
    theme(strip.text = element_text(size = 7))
}

fig_ba_unweighted <- ba_plot("Unweighted centroid")
ggsave("./tables_figures/fig_ba_unweighted_centroid.png", fig_ba_unweighted, width = 12, height = 10, dpi = 300)
fig_ba_unweighted

fig_ba_weighted <- ba_plot("Pop.-weighted centroid")
ggsave("./tables_figures/fig_ba_weighted_centroid.png", fig_ba_weighted, width = 12, height = 10, dpi = 300)
fig_ba_weighted
```

### Summary table: bias, LoA, and proportional bias

```{r bland-altman-table}
ba_summary %>%
  mutate(
    across(c(bias, bias_lo, bias_hi, loa_lo, loa_hi, slope), \(x) round(x, 3)),
    loa_width = round(loa_hi - loa_lo, 3),
    bias_fmt  = sprintf("%.3f [%.3f, %.3f]", bias, bias_lo, bias_hi),
    loa_fmt   = sprintf("[%.3f, %.3f]", loa_lo, loa_hi),
    p_fmt     = if_else(slope_p < 0.001, "<0.001", sprintf("%.3f", slope_p))
  ) %>%
  select(`Centroid` = centroid_label, `Retailer` = type, `Drive (min)` = drive,
         `N` = n, `Bias [95% CI]` = bias_fmt, `95% LoA` = loa_fmt,
         `LoA width` = loa_width, `B (slope)` = slope, `p (slope)` = p_fmt) %>%
  arrange(Centroid, Retailer, `Drive (min)`) %>%
  kable(caption = "Bland-Altman summary by centroid type, retailer, and drive-time threshold. Bias and LoA CIs use SE(bias)=s/\u221an and SE(LoA)=\u221a(3)s/\u221an with t_{n\u22121} (Bland & Altman 1999). \u03b2 = proportional-bias slope; * p<0.05.") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  row_spec(0, bold = TRUE) %>%
  collapse_rows(1:2, valign = "top")
```


