---
title: "2_prelim_results"
author: "Angela Zhang"
date: "2025-04-02"
output: html_document
---
# TODO: Summary of this file and mapping to aims: 
# NOTE: This is a standalone RMD file meaning it can run independently of other files in this codebase


```{r setup, include=TRUE}
source("0_Libraries.R")
library(kableExtra)
library(tidycensus)

library(Metrics)

# run summarize measures
# source('2_summarize_measures')
```

# Planning document
- [Methods description is here](https://docs.google.com/document/d/1rAyl0qsaWPcxM4E1X8-hP70vpzbPF7CYUYRq4pbERDM/edit?tab=t.0)


# Pull relevant spatial boundaries and datasets 

```{r setup}
la_city <- get_city_boundary(proj_crs) 

la_hh <- get_lac_households(proj_coord_crs)

la_ct <- get_census_tracts(proj_crs, state=proj_state, year=proj_year, county=proj_county) |>
  mutate(GEOID=as.character(GEOID)) %>%
  st_as_sf()

la_city_ct <- la_ct %>%
  dplyr::filter((lengths(st_intersects(., la_city)) > 0)) %>%
  st_transform(proj_coord_crs)

# include households with census tract in la city
la_city_hh <- la_hh %>%
  filter(GEOID_20 %in% la_city_ct$GEOID)

# get water
la_water <- area_water(state=proj_state, county=proj_county) %>%
  st_transform(proj_crs) %>%
  clipintersect_boundary(la_city)

la_openspace <- get_openspace_parks() |> 
    st_transform(proj_crs) 

# osm_parks <- getbb("Los Angeles County", display_name_contains = "United States") %>%
#   opq() %>% 
#   add_osm_feature(key = "leisure", value=c("park", "nature_reserve", "golf_course")) %>%
#   osmdata_sf()
# 
# la_parks <- osm_parks$osm_polygons %>%
#   st_transform(proj_crs) %>%
#   clipintersect_boundary(la_city)


```

# Sanity checks
✔️ We expect there to be 1424424 households in LA City
✔️ We also want to see why some CTs are missing observations in the household aggregate measures
  - No households were located in those CTs 
  - Want to check why that is (is the population zero? did they get misattributed to the wrong CTs?)

````{r}
# nrow(la_city_hh) # 1424424
# head(dt_household) 
# length(unique(dt_household$id)) # 1424424
```

### Load CT level data for all measures

```{r setup}

# TODO un-transform GEOID from as numeric 
# visualize ct_level data
# head(read.csv(paste0(processed_path, "/LAC_cleaned/ct_driving_times.csv"))) 
#'* Data frame with household values of food environment measures merged with census data*
parcel_data0 <- fread(paste0(processed_path, "LAC_cleaned/parcel_driving_all.csv"))

parcel_data_all <- parcel_data0 |> 
  tidytable::left_join(la_city_hh, by.x = c("id", "GEOID"), by.y = c("id", "GEOID_20"))


#'* Data frame with all food environment measures aggregated to the census tract level (geographic specificity) *
ct_drive_data <- la_ct %>%
  merge(read_csv(paste0(processed_path, "/LAC_cleaned/ct_driving_times.csv")), by= "GEOID")

#'* Data frame with parcel means of food environment measures only: used for demonstrating variation within census tracts *
parcel_data_ct <- read_csv(paste0(processed_path, "/LAC_cleaned/dt_household_ct.csv")) |>
  merge(la_ct, by= "GEOID")


temp <- read_csv(paste0(processed_path, "LAC_cleaned/dt_household_ct.csv"))

```

### (optional) Robustness check: Inspecting missing census tract data in the household dataset
```{r}
# missing <- setdiff(as.numeric(la_city_ct$GEOID), as.numeric(ct_drive_data$GEOID))
# missing2 <- setdiff(as.numeric(la_city_ct$GEOID), as.numeric(la_city_hh$GEOID_20))

# plot census tract data for parcels 
tmap_mode("plot")
map <-  tm_shape(la_city) +
  tm_polygons(col="yellow") + 
  tm_shape(la_city_ct[la_city_ct$GEOID %in% missing,] ) + 
  tm_polygons(col= "red", alpha=.5) +
  tm_shape(ct_drive_data) +
  tm_polygons(col = "driving_FF_15_parcel_mean", style = "quantile", palette = "Blues", colorNA = NULL)

# see if la households datasets are different from processed data
la_city_hhn <- la_hh %>%
  filter(lengths(st_intersects(., st_transform(la_city, proj_coord_crs))) > 0) 

la_city_hh <- la_hh %>%
  filter(GEOID_20 %in% la_city_ct$GEOID) %>%
  st_transform(proj_coord_crs)

missing2 <- setdiff(as.numeric(la_city_ct$GEOID), as.numeric(la_city_hh$GEOID_20))


#'*Figure out why NA values for ct_wtcent of a particular census tract * 
missing_wtct <- parcel_data_all %>% filter(is.na(ct_wtcent))
original_wtct <- dt_ct_wtcentm %>% filter(GEOID %in% missing_wtct$GEOID) 
# find that missing tract is population = 0 https://datacommons.org/place/geoId/06037980022
# should be CT: 106645 (misgeocoded)
data_to_conv <- parcel_data_all %>% filter(GEOID==6037980022)

```


# Data analysis

## Calculate metrics and agreements btwn parcel-level and census tract-level values
- Pearson's R (linear correlation)
- Spearman Rank p (non parametric correlation between variables)
- Bias
- MAE
- RMSE 
- TODO: figure out the source of missing data for weighted centroid
```{r}
sub <- parcel_data_all |> 
  filter(type=="SMK" & drive==15)

results <- parcel_data_all |>
  group_by(type, drive, network) |> 
  summarize(bias_hh_ct = bias(count, ct_cent), 
            mae_hh_ct = mae(count, ct_cent), 
            rmse_hh_ct = rmse(count, ct_cent),
            bias_hh_wct = bias(count, ct_wtcent),
            mae_hh_wct = mae(count, ct_wtcent), 
            rmse_hh_wct = rmse(count, ct_wtcent),
            bias_wct_ct = bias(ct_wtcent, ct_cent),
            mae_wct_ct = mae(ct_wtcent, ct_cent), 
            rmse_wct_ct = rmse(ct_wtcent, ct_cent))

data.table::fwrite(results, "./tables_figures/table3.csv")


```
## Plot histograms to observe distributions 
We find that many variables are not normally distributed so that leads us to select the Wilcoxon rank sum test to test differences in means between the various groups

*Todo* Additionally, we want to try to identify what type of measurement error/bias this could lead to which requires observing the distributions of the parcel-level values compared to the aggregate values (among other techniques). The figures below are a start

*Todo* look into the zero values for CT_CENT that are shown in the third figure
```{r}
dist <- ct_drive_data |>
  as.data.table() |>
  melt(measure.vars=c("ct_cent", "ct_wtcent", "parcel_mean")) 

ggplot(dist, aes(x=value, fill=variable, alpha=.5)) + 
   geom_step(stat = "bin", bins = 50, direction = "mid") +
  geom_histogram(bins = 50, alpha = 0.3, colour = NA,
    position = "identity") +
  facet_grid(cols=vars(type), rows=vars(drive), scales="free") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

ggplot(dist |> filter(type=="SMK"), aes(x=value, col=variable, fill=variable, alpha=.3)) +
  geom_step(stat = "bin", bins = 50, direction = "mid") +
  geom_histogram(bins = 50, alpha = 0.3, colour = NA,
    position = "identity") +
  #geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  facet_grid(rows=vars(drive), scales="free") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

```

## Distribution of parcel, centroid, and weighted centroid measures (faceted by drive time × retailer)
```{r dist-all-measures, fig.width=14, fig.height=10}
measure_labels <- c(
  parcel_mean = "Parcel (ground truth)",
  ct_cent     = "CT centroid",
  ct_wtcent   = "Pop.-weighted centroid"
)

dist_long <- ct_drive_data |>
  st_drop_geometry() |>
  filter(type %in% c("CNV", "FF", "GRC", "RR", "SMK", "SPF")) |>
  pivot_longer(c(parcel_mean, ct_cent, ct_wtcent),
               names_to = "measure", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(
    measure = factor(measure,
                     levels = c("parcel_mean", "ct_cent", "ct_wtcent"),
                     labels = unname(measure_labels)),
    drive_label = paste0(drive, " min")
  )

ggplot(dist_long, aes(x = value, colour = measure, fill = measure)) +
  geom_density(alpha = 0.20, linewidth = 0.6) +
  facet_wrap(~drive_label+type,
    scales = "free"
  ) +
  scale_colour_manual(values = c("Parcel (ground truth)" = "#2166ac",
                                 "CT centroid"           = "#d6604d",
                                 "Pop.-weighted centroid"= "#4dac26")) +
  scale_fill_manual(  values = c("Parcel (ground truth)" = "#2166ac",
                                 "CT centroid"           = "#d6604d",
                                 "Pop.-weighted centroid"= "#4dac26")) +
  labs(
    x      = "Accessibility (count within threshold)",
    y      = "Density",
    colour = "Measure",
    fill   = "Measure",
    title  = "Distribution of accessibility measures by drive time and retailer type",
    subtitle = "Rows = drive-time threshold (min); Columns = retailer type"
  ) +
  theme_bw(base_size = 9) +
  theme(
    legend.position   = "bottom",
    strip.text        = element_text(size = 7, face = "bold"),
    axis.text.x       = element_text(angle = 45, vjust = 1, hjust = 1),
    panel.grid.minor  = element_blank()
  )
```
## Calculate t-tests to evaluate differences in distributions for census tract metrics and parcel metrics aggregated to census level 
```{r}
i=1
ttests0 <- split(ct_drive_data, list(ct_drive_data$type, ct_drive_data$drive)) %>%
  lapply(
  function(df) { 
    i <<- i+1
    print(length(df$parcel_mean))
      # paired n for each comparison (drops NAs pairwise)
    n1 <- sum(complete.cases(df$parcel_mean, df$ct_cent))
    n2 <- sum(complete.cases(df$ct_wtcent, df$ct_cent))
    n3 <- sum(complete.cases(df$parcel_mean, df$ct_wtcent))
    
    t1 <<- t.test(df$parcel_mean, df$ct_cent, paired=TRUE, alternative="two.sided")
    t2 <- t.test(df$ct_wtcent, df$ct_cent, paired=TRUE, alternative="two.sided")
    t3 <- t.test(df$parcel_mean, df$ct_wtcent, paired =TRUE, alternative="two.sided")

    
    drive <- unique(df$drive)[1]
    print(drive)
    type <- unique(df$type)[1]
    print(type)
    res <- data.frame(
      time_dist = unique(df$drive)[1], 
      type =  unique(df$type)[1],
      tests = c("parcel_cent", "wtcent_cent", "parcel_wtcent"), 
      n = c(n1, n2, n3),
      pval = c(t1$p.value, t2$p.value, t3$p.value),
      est = c(t1$estimate, t2$estimate, t3$estimate),
         conf.low  = c(t1$conf.int[1], t2$conf.int[1], t3$conf.int[1]),
      conf.high = c(t1$conf.int[2], t2$conf.int[2], t3$conf.int[2])
    )
    print(length(res))
    return(res)
  }
)

ttests <- ttests0 |>
  bind_rows()

ttests %>%
  select(type, time_dist, tests, n, est, conf.low, conf.high, pval) %>%
  kable(digits = 4, caption = "Paired t-test results") %>%
    row_spec(0, bold = TRUE, background = "white", color = "black") 
```
## Plotting results of t-test 
The results of the t-test shows that for longer drive time distances (>25 minutes), means of the census-tract level parcel measure are significantly different from means of the centroid calculated measure and means of population-centroid calculated measures 
```{r}
ggplot(ttests, aes(x=fct_relevel(tests, "parcel_cent", "wtcent_cent", "parcel_wtcent"), y=est, color=(pval<.05))) + 
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  facet_grid(cols=vars(time_dist), rows=vars(type), scales="free") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

ggplot(ct_drive_data, aes(x=parcel_mean, y=ct_wtcent)) + 
  geom_point(size=1, aes(alpha=.5)) + 
  #geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  facet_grid(cols=vars(drive), rows=vars(type), scales="fixed") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

unique(ct_drive_data$type)

# TODO make table of the t-tests
# TODO make figure of t-tests

```
## Wilcox test results
Given that distributions of counts are not strictly following a normal distribution, we also perform a Wilcox test
```{r}
wtests0 <- split(ct_drive_data, list(ct_drive_data$type, ct_drive_data$drive)) %>%
  lapply(
  function(df) { 
   # i <<- i+1
    print(length(df$parcel_mean))
      # paired n for each comparison (drops NAs pairwise)
    n1 <- sum(complete.cases(df$parcel_mean, df$ct_cent))
    n2 <- sum(complete.cases(df$ct_wtcent, df$ct_cent))
    n3 <- sum(complete.cases(df$parcel_mean, df$ct_wtcent))

    
    t1 <<- wilcox.test(df$parcel_mean, df$ct_cent, paired=TRUE, alternative="two.sided")
    t2 <- wilcox.test(df$ct_wtcent, df$ct_cent, paired=TRUE, alternative="two.sided")
    t3 <- wilcox.test(df$parcel_mean, df$ct_wtcent, paired =TRUE, alternative="two.sided")

    drive <- unique(df$drive)[1]
    print(drive)
    type <- unique(df$type)[1]
    print(type)
    res <- data.frame(
      time_dist = unique(df$drive)[1], 
      type =  unique(df$type)[1],
      tests = c("parcel_cent", "wtcent_cent", "parcel_wtcent"), 
      n = c(n1, n2, n3),
      pval = c(t1$p.value, t2$p.value, t3$p.value)
      #est = c(t1$estimate, t2$estimate, t3$estimate),
      #conf.low  = c(t1$conf.int[1], t2$conf.int[1], t3$conf.int[1]),
      #conf.high = c(t1$conf.int[2], t2$conf.int[2], t3$conf.int[2])
    )
    print(length(res))
    return(res)
  }
)

wtests <- wtests0 |>
  bind_rows() |> 
  pivot_wider(values_from=pval, names_from=c(tests,n)) %>%
  select(type, time_dist, everything())%>%
  arrange(type)

wtests %>%
  #select(type, time_dist, tests, n, pval) %>%
  kable(digits = 4, caption = "Paired Wilcox (Mann Whitney U)-test results", booktabs = TRUE) %>%
  row_spec(0, bold = TRUE, background = "white", color = "black") |>
  kableExtra::collapse_rows(valign = "top")

unique(ct_drive_data$type)

```

# Mapping data 
## Map variation within CTs using parcel data
- map parcel data for one rural tract, one urban tract, one suburban tract
*todo* figure out why there are missing values for 5 minutes
- map coefficient of variation (dispersion) for each CT in LA City using parcel data
```{r}
tm_shape(parcel_data_ct %>% drop_na(driving_SMK_5_parcel_cv)) + 
  tm_polygons(col="driving_SMK_5_parcel_cv", style="jenks") + 
  tm_shape(la_openspace) +
  tm_fill(col="green", lwd=0) +  
  tm_shape(la_water) +
  tm_fill(col="lightblue", lwd=0) 


tm_shape(parcel_data_ct %>% drop_na(driving_SMK_5_parcel_cv)) + 
  tm_polygons(col="driving_SMK_10_parcel_cv", style="jenks") +  
  tm_shape(la_openspace) +
  tm_fill(col="green", lwd=0) +  
  tm_shape(la_water) +
  tm_fill(col="lightblue", lwd=0) 

tm_shape(parcel_data_ct %>% drop_na(driving_SMK_5_parcel_cv)) + 
  tm_polygons(col="driving_SMK_15_parcel_cv", style="jenks") + 
  tm_shape(la_openspace) +
  tm_fill(col="green", lwd=0) +  
  tm_shape(la_water) +
  tm_fill(col="lightblue", lwd=0) 

tm_shape(parcel_data_ct %>% drop_na(driving_SMK_5_parcel_cv)) + 
  tm_polygons(col="driving_SMK_20_parcel_cv", style="jenks", legend.show = FALSE) + 
  tm_shape(la_openspace) +
  tm_fill(col="green", lwd=0) +  
  tm_shape(la_water) +
  tm_fill(col="lightblue", lwd=0) + 
  tm_scale_bar(position=c("LEFT", "bottom"))
```
 
## Table 4: Census-tract features by quartiles of coefficient of variation
Pull ACS demographic/socioeconomic features and summarize by CV quartile. CV is based on the within-CT dispersion of parcel-level supermarket accessibility (15-min drive).

```{r table4-cv-quartiles}
library(tidycensus)
# TODO missing CVs in some CTs (show up as NA) 06037300301

# ACS variables: population, income, poverty, education, Hispanic share
acs_vars <- c(
  "B01003_001", # Total population
  "B19013_001", # Median household income
  "B17001_001", # Poverty universe
  "B17001_002", # Below poverty level
  "B15003_001", # Population 25+ (education denominator)
  "B15003_022", # Bachelor's degree
  "B15003_023", # Master's degree
  "B15003_024", # Professional school degree
  "B15003_025", # Doctorate degree
  "B03003_001", # Total population (Hispanic origin table)
  "B03003_003"  # Hispanic or Latino
)

ct_acs <- get_acs(
  geography = "tract",
  variables = acs_vars,
  state = proj_state,
  year = proj_year,
  output = "wide",
  geometry = FALSE
) %>%
  transmute(
    GEOID        = GEOID,
    pop_total    = B01003_001E,
    hhinc_median = B19013_001E,
    pct_poverty  = B17001_002E / B17001_001E * 100,
    pct_college  = (B15003_022E + B15003_023E + B15003_024E + B15003_025E) / B15003_001E * 100,
    pct_hispanic = B03003_003E / B03003_001E * 100
  )

# Land/water area from tigris (ALAND, AWATER columns present in la_ct)
ct_area <- la_ct %>%
  st_drop_geometry() %>%
  transmute(
    GEOID      = as.character(GEOID),
    aland_km2  = ALAND / 1e6,
    awater_km2 = AWATER / 1e6
  )

# Formatting helpers (defined once, used inside make_table4)
fmt_meansd  <- function(x) sprintf("%.1f ± %.1f", mean(x, na.rm = TRUE), sd(x, na.rm = TRUE))
fmt_med_iqr <- function(x) sprintf("%.0f [%.0f–%.0f]",
  median(x, na.rm = TRUE), quantile(x, .25, na.rm = TRUE), quantile(x, .75, na.rm = TRUE))
fmt_pval    <- function(p) if (p < 0.001) "<0.001" else sprintf("%.3f", p)

# Build and display Table 4 for one retailer type × drive-time combination
make_table4 <- function(type, drive, parcel_data_ct, ct_acs, ct_area) {
  cv_col <- paste0("driving_", type, "_", drive, "_parcel_cv")
  if (!cv_col %in% names(parcel_data_ct)) return(NULL)

  ct_cv_tbl <- parcel_data_ct %>%
    st_drop_geometry() %>%
    select(GEOID, all_of(cv_col)) %>%
    rename(cv_value = all_of(cv_col)) %>%
    left_join(ct_acs, by = "GEOID") %>%
    left_join(ct_area, by = "GEOID") %>%
    mutate(
      pop_density = pop_total / aland_km2,
      cv_quartile = ntile(cv_value, 4),
      cv_quartile = factor(cv_quartile, labels = c("Q1 (Low CV)", "Q2", "Q3", "Q4 (High CV)"))
    )

  table4 <- ct_cv_tbl %>%
    group_by(cv_quartile) %>%
    summarise(
      `N tracts`                             = n(),
      `CV, mean ± SD`                        = fmt_meansd(cv_value),
      `Population, median [IQR]`             = fmt_med_iqr(pop_total),
      `Pop. density (per km²), median [IQR]` = fmt_med_iqr(pop_density),
      `Land area (km²), median [IQR]`        = sprintf("%.2f [%.2f–%.2f]",
          median(aland_km2, na.rm = TRUE),
          quantile(aland_km2, .25, na.rm = TRUE),
          quantile(aland_km2, .75, na.rm = TRUE)),
      `% Hispanic, mean ± SD`                = fmt_meansd(pct_hispanic),
      `% Below poverty, mean ± SD`           = fmt_meansd(pct_poverty),
      `% College educated, mean ± SD`        = fmt_meansd(pct_college),
      `Median HH income, median [IQR]`       = sprintf("$%.0f [$%.0f–$%.0f]",
          median(hhinc_median, na.rm = TRUE),
          quantile(hhinc_median, .25, na.rm = TRUE),
          quantile(hhinc_median, .75, na.rm = TRUE))
    ) %>%
    rename(`CV Quartile` = cv_quartile)

  kw_pvals <- tibble::tribble(
    ~Variable,                                ~`p-value`,
    "N tracts",                               NA_character_,
    "CV, mean ± SD",                          fmt_pval(kruskal.test(cv_value     ~ cv_quartile, ct_cv_tbl)$p.value),
    "Population, median [IQR]",               fmt_pval(kruskal.test(pop_total    ~ cv_quartile, ct_cv_tbl)$p.value),
    "Pop. density (per km²), median [IQR]",   fmt_pval(kruskal.test(pop_density  ~ cv_quartile, ct_cv_tbl)$p.value),
    "Land area (km²), median [IQR]",          fmt_pval(kruskal.test(aland_km2    ~ cv_quartile, ct_cv_tbl)$p.value),
    "% Hispanic, mean ± SD",                  fmt_pval(kruskal.test(pct_hispanic ~ cv_quartile, ct_cv_tbl)$p.value),
    "% Below poverty, mean ± SD",             fmt_pval(kruskal.test(pct_poverty  ~ cv_quartile, ct_cv_tbl)$p.value),
    "% College educated, mean ± SD",          fmt_pval(kruskal.test(pct_college  ~ cv_quartile, ct_cv_tbl)$p.value),
    "Median HH income, median [IQR]",         fmt_pval(kruskal.test(hhinc_median ~ cv_quartile, ct_cv_tbl)$p.value)
  )

  table4 %>%
    mutate(`N tracts` = as.character(`N tracts`)) %>%
    pivot_longer(-`CV Quartile`, names_to = "Variable", values_to = "val") %>%
    pivot_wider(names_from = `CV Quartile`, values_from = val) %>%
    left_join(kw_pvals, by = "Variable") %>%
    select(Variable, everything())
}

# Detect all available type × drive combinations from parcel_data_ct column names
cv_cols <- grep("^driving_.+_\\d+_parcel_cv$", names(parcel_data_ct), value = TRUE)
combos  <- data.frame(
  type  = sub("^driving_(.+)_(\\d+)_parcel_cv$", "\\1", cv_cols),
  drive = as.integer(sub("^driving_(.+)_(\\d+)_parcel_cv$", "\\2", cv_cols))
) %>% arrange(type, drive)

# Generate, display, and save Table 4 for every combination
for (i in seq_len(nrow(combos))) {
  type  <- combos$type[i]
  drive <- combos$drive[i]
  tbl   <- make_table4(type, drive, parcel_data_ct, ct_acs, ct_area)
  if (is.null(tbl)) next

  caption_str <- sprintf(
    "Table 4. CT features by CV quartile (%s, %d-min drive). p from Kruskal-Wallis.",
    type, drive
  )
  print(
    tbl %>%
      kable(caption = caption_str,
            align   = c("l", rep("r", ncol(tbl) - 1))) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
      row_spec(0, bold = TRUE)
  )

  data.table::fwrite(tbl, sprintf("./tables_figures/table4_%s_%d.csv", type, drive))
}
```

## Bland-Altman Analysis: Centroid vs. Parcel Agreement

```{r bland-altman, fig.width=12, fig.height=10}
library(blandr)

# Long-form: one row per CT × retailer × drive × centroid type
ba_data <- ct_drive_data %>%
  st_drop_geometry() %>%
  filter(!is.na(parcel_mean)) %>%
  pivot_longer(c(ct_cent, ct_wtcent), names_to = "centroid_type", values_to = "centroid") %>%
  filter(!is.na(centroid)) %>%
  mutate(
    diff = centroid - parcel_mean,
    mean = (centroid + parcel_mean) / 2,
    centroid_label = factor(centroid_type,
      levels = c("ct_cent", "ct_wtcent"),
      labels = c("Unweighted centroid", "Pop.-weighted centroid"))
  )

# BA statistics + proportional bias per group (one blandr call per group)
# CIs follow Bland & Altman (1986/1999): SE(bias) = s/sqrt(n), SE(LoA) = sqrt(3)*s/sqrt(n),
# using t-distribution with n-1 df — as implemented in blandr.statistics()
ba_summary <- ba_data %>%
  group_by(centroid_label, type, drive) %>%
  group_modify(~ {
    s   <- blandr.statistics(.x$centroid, .x$parcel_mean)
    fit <- lm(diff ~ mean, data = .x)
    tibble(
      n          = nrow(.x),
      bias       = s$bias,       bias_lo  = s$biasLowerCI,      bias_hi  = s$biasUpperCI,
      loa_lo     = s$lowerLOA,   loa_lo_lo = s$lowerLOA_lowerCI, loa_lo_hi = s$lowerLOA_upperCI,
      loa_hi     = s$upperLOA,   loa_hi_lo = s$upperLOA_lowerCI, loa_hi_hi = s$upperLOA_upperCI,
      slope      = coef(fit)["mean"],
      slope_p    = coef(summary(fit))["mean", "Pr(>|t|)"]
    )
  }) %>%
  mutate(slope_label = sprintf("\u03b2=%.3f%s", slope, if_else(slope_p < 0.05, "*", "")))

# Helper: BA plot for one centroid type, faceted by retailer × drive time
# CI bands on bias and LoA lines per Bland & Altman (1999)
ba_plot <- function(centroid_lbl) {
  d    <- filter(ba_data,    centroid_label == centroid_lbl)
  smry <- filter(ba_summary, centroid_label == centroid_lbl)

  ggplot(d, aes(mean, diff)) +
    geom_point(alpha = 0.2, size = 0.6, colour = "grey40") +
    # 95% CI bands for bias and LoA (shaded rectangles spanning full x range)
    geom_rect(data = smry, aes(xmin=-Inf, xmax=Inf, ymin=bias_lo,  ymax=bias_hi),
              fill = "steelblue", alpha = 0.15, inherit.aes = FALSE) +
    geom_rect(data = smry, aes(xmin=-Inf, xmax=Inf, ymin=loa_lo_lo, ymax=loa_lo_hi),
              fill = "tomato",    alpha = 0.15, inherit.aes = FALSE) +
    geom_rect(data = smry, aes(xmin=-Inf, xmax=Inf, ymin=loa_hi_lo, ymax=loa_hi_hi),
              fill = "tomato",    alpha = 0.15, inherit.aes = FALSE) +
    # Bias and LoA lines
    geom_hline(data = smry, aes(yintercept = bias),   colour = "steelblue", linewidth = 0.8) +
    geom_hline(data = smry, aes(yintercept = loa_lo), colour = "tomato", linetype = "dashed") +
    geom_hline(data = smry, aes(yintercept = loa_hi), colour = "tomato", linetype = "dashed") +
    # Proportional bias regression line
    geom_smooth(method = "lm", se = TRUE, colour = "darkorange", linewidth = 0.6, alpha = 0.15) +
    geom_text(data = smry, aes(x = -Inf, y = Inf, label = slope_label),
              hjust = -0.1, vjust = 1.5, size = 2.5, colour = "darkorange4") +
    geom_hline(yintercept = 0, linewidth = 0.3) +
    facet_grid(type ~ drive, labeller = labeller(drive = \(x) paste0(x, " min")), scales = "free") +
    labs(x = "Mean of centroid and parcel", y = "Centroid \u2212 Parcel",
         title = paste("Bland-Altman:", centroid_lbl, "vs. parcel (ground truth)"),
         subtitle = "Blue = bias \u00b1 95% CI  |  Red dashed = 95% LoA \u00b1 95% CI  |  Orange = proportional-bias fit (* p<0.05)") +
    theme_bw(base_size = 9) +
    theme(strip.text = element_text(size = 7))
}

ba_plot("Unweighted centroid")
ba_plot("Pop.-weighted centroid")
```

### Summary table: bias, LoA, and proportional bias

```{r bland-altman-table}
ba_summary %>%
  mutate(
    across(c(bias, bias_lo, bias_hi, loa_lo, loa_hi, slope), \(x) round(x, 3)),
    loa_width = round(loa_hi - loa_lo, 3),
    bias_fmt  = sprintf("%.3f [%.3f, %.3f]", bias, bias_lo, bias_hi),
    loa_fmt   = sprintf("[%.3f, %.3f]", loa_lo, loa_hi),
    p_fmt     = if_else(slope_p < 0.001, "<0.001", sprintf("%.3f", slope_p))
  ) %>%
  select(`Centroid` = centroid_label, `Retailer` = type, `Drive (min)` = drive,
         `N` = n, `Bias [95% CI]` = bias_fmt, `95% LoA` = loa_fmt,
         `LoA width` = loa_width, `B (slope)` = slope, `p (slope)` = p_fmt) %>%
  arrange(Centroid, Retailer, `Drive (min)`) %>%
  kable(caption = "Bland-Altman summary by centroid type, retailer, and drive-time threshold. Bias and LoA CIs use SE(bias)=s/\u221an and SE(LoA)=\u221a(3)s/\u221an with t_{n\u22121} (Bland & Altman 1999). \u03b2 = proportional-bias slope; * p<0.05.") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  row_spec(0, bold = TRUE) %>%
  collapse_rows(1:2, valign = "top")
```

## Compare to health outcomes (need to generate evidence-based measures)
*TODO* 
- Urban vs Rural varied buffer time (e.g how the USDA does it) 
- Use ratio measures 
- Include USDA food access measures as a way to test importance 

## Map El Sendero data 
- need to link to geocoded data (ask beau for this?) 
```{r}
# elsend <- read.csv("../../0_shared-data/latino-health-elsendero/raw/ElSenderoPathwaysToH_DATA_LABELS_2025-04-22_1153.csv")
# elsend1 <- read.csv("../../0_shared-data/latino-health-elsendero/raw/ElSenderoPathwaysToH_DATA_LABELS_2025-04-22_1154.csv")

# map the results

```

