---
title: "2_prelim_results"
author: "Angela Zhang"
date: "2025-04-02"
output: html_document
---
# TODO: Summary of this file and mapping to aims: 
# NOTE: This is a standalone RMD file meaning it can run independently of other files in this codebase


```{r setup, include=FALSE}
source("0_Libraries.R")
library(kableExtra)
library(Metrics)

# run summarize measures
# source('2_summarize_measures')
```

# Planning document
- [Methods description is here](https://docs.google.com/document/d/1rAyl0qsaWPcxM4E1X8-hP70vpzbPF7CYUYRq4pbERDM/edit?tab=t.0)

# Sanity checks
✔️ We expect there to be 1424424 households in LA City
✔️ We also want to see why some CTs are missing observations in the household aggregate measures
  - No households were located in those CTs 
  - Want to check why that is (is the population zero? did they get misattributed to the wrong CTs?)

````{r}
nrow(la_city_hh) # 1424424
head(dt_household) 
length(unique(dt_household$id)) # 1424424
```

# Pull relevant spatial boundaries and datasets 

```{r}
la_city <- get_city_boundary(proj_crs) 

la_hh <- get_lac_households(4326)

la_ct <- get_census_tracts(proj_crs, state="CA", year=2020, county="Los Angeles") |>
  mutate(GEOID=as.numeric(GEOID)) %>%
  st_as_sf() 

la_city_ct <- la_ct %>%
  dplyr::filter((lengths(st_intersects(., la_city)) > 0)) %>%
  st_transform(4326)

# include households with census tract in la city
la_city_hh <- la_hh %>%
  filter(as.numeric(GEOID_20) %in% as.numeric(la_city_ct$GEOID)) 

rm(la_hh)

# get water
la_water <- area_water(state="CA", county="Los Angeles") %>%
  st_transform(proj_crs) %>%
  clipintersect_boundary(la_city)

la_openspace <- get_openspace_parks() |> 
    st_transform(proj_crs) 

# osm_parks <- getbb("Los Angeles County", display_name_contains = "United States") %>%
#   opq() %>% 
#   add_osm_feature(key = "leisure", value=c("park", "nature_reserve", "golf_course")) %>%
#   osmdata_sf()
# 
# la_parks <- osm_parks$osm_polygons %>%
#   st_transform(proj_crs) %>%
#   clipintersect_boundary(la_city)


```

### Load CT level data for all measures

```{r cars}

# TODO un-transform GEOID from as numeric 
# visualize ct_level data
# head(read.csv(paste0(processed_path, "/LAC_cleaned/ct_driving_times.csv"))) 
#'* Data frame with household values of food environment measures merged with census data*
parcel_data0 <- fread(paste0(processed_path, "LAC_cleaned/parcel_driving_all.csv"))

parcel_data_all <- parcel_data0 |> 
  tidytable::left_join(la_city_hh, by.x = c("id", "GEOID"), by.y = c("id", "GEOID_20"))


#'* Data frame with all food environment measures aggregated to the census tract level (geographic specificity) *
ct_drive_data <- la_ct %>%
  merge(read_csv(paste0(processed_path, "/LAC_cleaned/ct_driving_times.csv")), by= "GEOID")

#'* Data frame with parcel means of food environment measures only: used for demonstrating variation within census tracts *
parcel_data_ct <- la_ct %>%
  merge(read_csv(paste0(processed_path, "/LAC_cleaned/dt_household_ct.csv")), by= "GEOID")


temp <- read_csv(paste0(processed_path, "LAC_cleaned/dt_household_ct.csv"))

```

### (optional) Robustness check: Inspecting missing census tract data in the household dataset
```{r}
# missing <- setdiff(as.numeric(la_city_ct$GEOID), as.numeric(ct_drive_data$GEOID))
# missing2 <- setdiff(as.numeric(la_city_ct$GEOID), as.numeric(la_city_hh$GEOID_20))

# plot census tract data for parcels 
tmap_mode("plot")
map <-  tm_shape(la_city) +
  tm_polygons(col="yellow") + 
  tm_shape(la_city_ct[la_city_ct$GEOID %in% missing,] ) + 
  tm_polygons(col= "red", alpha=.5) +
  tm_shape(ct_drive_data) +
  tm_polygons(col = "driving_FF_15_parcel_mean", style = "quantile", palette = "Blues", colorNA = NULL)

# see if la households datasets are different from processed data
la_city_hhn <- la_hh %>%
  filter(lengths(st_intersects(., st_transform(la_city, 4326))) > 0) 

la_city_hh <- la_hh %>%
  filter(GEOID_20 %in% la_city_ct$GEOID) %>%
  st_transform(4326)

missing2 <- setdiff(as.numeric(la_city_ct$GEOID), as.numeric(la_city_hh$GEOID_20))


#'*Figure out why NA values for ct_wtcent of a particular census tract * 
missing_wtct <- parcel_data_all %>% filter(is.na(ct_wtcent))
original_wtct <- dt_ct_wtcentm %>% filter(GEOID %in% missing_wtct$GEOID) 
# find that missing tract is population = 0 https://datacommons.org/place/geoId/06037980022
# should be CT: 106645 (misgeocoded)
data_to_conv <- parcel_data_all %>% filter(GEOID==6037980022)

print(map)
```


# Data analysis

## Calculate metrics and agreements btwn parcel-level and census tract-level values
- Pearson's R (linear correlation)
- Spearman Rank p (non parametric correlation between variables)
- Bias
- MAE
- RMSE 
- TODO: figure out the source of missing data for weighted centroid
```{r}
sub <- parcel_data_all |> 
  filter(type=="SMK" & drive==15)

results <- parcel_data_all |>
  group_by(type, drive, network) |> 
  summarize(bias_hh_ct = bias(count, ct_cent), 
            mae_hh_ct = mae(count, ct_cent), 
            rmse_hh_ct = rmse(count, ct_cent),
            bias_hh_wct = bias(count, ct_wtcent),
            mae_hh_wct = mae(count, ct_wtcent), 
            rmse_hh_wct = rmse(count, ct_wtcent),
            bias_wct_ct = bias(ct_wtcent, ct_cent),
            mae_wct_ct = mae(ct_wtcent, ct_cent), 
            rmse_wct_ct = rmse(ct_wtcent, ct_cent))
print(results)

data.table::fwrite(results, "./tables_figures/table3.csv")


```
## Plot histograms to observe distributions 
We find that many variables are not normally distributed so that leads us to select the Wilcoxon rank sum test to test differences in means between the various groups

*Todo* Additionally, we want to try to identify what type of measurement error/bias this could lead to which requires observing the distributions of the parcel-level values compared to the aggregate values (among other techniques). The figures below are a start

*Todo* look into the zero values for CT_CENT that are shown in the third figure
```{r}
dist <- ct_drive_data |>
  as.data.table() |>
  melt(measure.vars=c("ct_cent", "ct_wtcent", "parcel_mean")) 

ggplot(dist, aes(x=value, fill=variable, alpha=.5)) + 
   geom_step(stat = "bin", bins = 50, direction = "mid") +
  geom_histogram(bins = 50, alpha = 0.3, colour = NA,
    position = "identity") +
  facet_grid(cols=vars(type), rows=vars(drive), scales="free") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

ggplot(dist |> filter(type=="SMK"), aes(x=value, col=variable, fill=variable, alpha=.3)) + 
  geom_step(stat = "bin", bins = 50, direction = "mid") +
  geom_histogram(bins = 50, alpha = 0.3, colour = NA,
    position = "identity") +
  #geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  facet_grid(rows=vars(drive), scales="free") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

```
## Calculate t-tests to evaluate differences in distributions for census tract metrics and parcel metrics aggregated to census level 
```{r}
i=1
ttests0 <- split(ct_drive_data, list(ct_drive_data$type, ct_drive_data$drive)) %>%
  lapply(
  function(df) { 
    i <<- i+1
    print(length(df$parcel_mean))
      # paired n for each comparison (drops NAs pairwise)
    n1 <- sum(complete.cases(df$parcel_mean, df$ct_cent))
    n2 <- sum(complete.cases(df$ct_wtcent, df$ct_cent))
    n3 <- sum(complete.cases(df$parcel_mean, df$ct_wtcent))
    
    t1 <<- t.test(df$parcel_mean, df$ct_cent, paired=TRUE, alternative="two.sided")
    t2 <- t.test(df$ct_wtcent, df$ct_cent, paired=TRUE, alternative="two.sided")
    t3 <- t.test(df$parcel_mean, df$ct_wtcent, paired =TRUE, alternative="two.sided")

    
    drive <- unique(df$drive)[1]
    print(drive)
    type <- unique(df$type)[1]
    print(type)
    res <- data.frame(
      time_dist = unique(df$drive)[1], 
      type =  unique(df$type)[1],
      tests = c("parcel_cent", "wtcent_cent", "parcel_wtcent"), 
      n = c(n1, n2, n3),
      pval = c(t1$p.value, t2$p.value, t3$p.value),
      est = c(t1$estimate, t2$estimate, t3$estimate),
         conf.low  = c(t1$conf.int[1], t2$conf.int[1], t3$conf.int[1]),
      conf.high = c(t1$conf.int[2], t2$conf.int[2], t3$conf.int[2])
    )
    print(length(res))
    return(res)
  }
)

ttests <- ttests0 |>
  bind_rows()

ttests %>%
  select(type, time_dist, tests, n, est, conf.low, conf.high, pval) %>%
  kable(digits = 4, caption = "Paired t-test results") %>%
    row_spec(0, bold = TRUE, background = "white", color = "black") 
```
## Plotting results of t-test 
The results of the t-test shows that for longer drive time distances (>25 minutes), means of the census-tract level parcel measure are significantly different from means of the centroid calculated measure and means of population-centroid calculated measures 
```{r}
ggplot(ttests, aes(x=fct_relevel(tests, "parcel_cent", "wtcent_cent", "parcel_wtcent"), y=est, color=(pval<.05))) + 
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  facet_grid(cols=vars(time_dist), rows=vars(type), scales="free") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

ggplot(ct_drive_data, aes(x=parcel_mean, y=ct_wtcent)) + 
  geom_point(size=1, aes(alpha=.5)) + 
  #geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  facet_grid(cols=vars(drive), rows=vars(type), scales="fixed") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

unique(ct_drive_data$type)

# TODO make table of the t-tests
# TODO make figure of t-tests

```
## Wilcox test results
Given that distributions of counts are not strictly following a normal distribution, we also perform a Wilcox test
```{r}
wtests0 <- split(ct_drive_data, list(ct_drive_data$type, ct_drive_data$drive)) %>%
  lapply(
  function(df) { 
   # i <<- i+1
    print(length(df$parcel_mean))
      # paired n for each comparison (drops NAs pairwise)
    n1 <- sum(complete.cases(df$parcel_mean, df$ct_cent))
    n2 <- sum(complete.cases(df$ct_wtcent, df$ct_cent))
    n3 <- sum(complete.cases(df$parcel_mean, df$ct_wtcent))

    
    t1 <<- wilcox.test(df$parcel_mean, df$ct_cent, paired=TRUE, alternative="two.sided")
    t2 <- wilcox.test(df$ct_wtcent, df$ct_cent, paired=TRUE, alternative="two.sided")
    t3 <- wilcox.test(df$parcel_mean, df$ct_wtcent, paired =TRUE, alternative="two.sided")

    drive <- unique(df$drive)[1]
    print(drive)
    type <- unique(df$type)[1]
    print(type)
    res <- data.frame(
      time_dist = unique(df$drive)[1], 
      type =  unique(df$type)[1],
      tests = c("parcel_cent", "wtcent_cent", "parcel_wtcent"), 
      n = c(n1, n2, n3),
      pval = c(t1$p.value, t2$p.value, t3$p.value)
      #est = c(t1$estimate, t2$estimate, t3$estimate),
      #conf.low  = c(t1$conf.int[1], t2$conf.int[1], t3$conf.int[1]),
      #conf.high = c(t1$conf.int[2], t2$conf.int[2], t3$conf.int[2])
    )
    print(length(res))
    return(res)
  }
)

wtests <- wtests0 |>
  bind_rows() |> 
  pivot_wider(values_from=pval, names_from=c(tests,n)) %>%
  select(type, time_dist, everything())%>%
  arrange(type)

wtests %>%
  #select(type, time_dist, tests, n, pval) %>%
  kable(digits = 4, caption = "Paired Wilcox (Mann Whitney U)-test results", booktabs = TRUE) %>%
  row_spec(0, bold = TRUE, background = "white", color = "black") |>
  kableExtra::collapse_rows(valign = "top")

unique(ct_drive_data$type)

```

# Mapping data 
## Map variation within CTs using parcel data
- map parcel data for one rural tract, one urban tract, one suburban tract
*todo* figure out why there are missing values for 5 minutes
- map coefficient of variation (dispersion) for each CT in LA City using parcel data
```{r}
tm_shape(parcel_data_ct %>% drop_na(driving_SMK_5_parcel_cv)) + 
  tm_polygons(col="driving_SMK_5_parcel_cv", style="jenks") + 
  tm_shape(la_openspace) +
  tm_fill(col="green", lwd=0) +  
  tm_shape(la_water) +
  tm_fill(col="lightblue", lwd=0) 


tm_shape(parcel_data_ct %>% drop_na(driving_SMK_5_parcel_cv)) + 
  tm_polygons(col="driving_SMK_10_parcel_cv", style="jenks") +  
  tm_shape(la_openspace) +
  tm_fill(col="green", lwd=0) +  
  tm_shape(la_water) +
  tm_fill(col="lightblue", lwd=0) 

tm_shape(parcel_data_ct %>% drop_na(driving_SMK_5_parcel_cv)) + 
  tm_polygons(col="driving_SMK_15_parcel_cv", style="jenks") + 
  tm_shape(la_openspace) +
  tm_fill(col="green", lwd=0) +  
  tm_shape(la_water) +
  tm_fill(col="lightblue", lwd=0) 

tm_shape(parcel_data_ct %>% drop_na(driving_SMK_5_parcel_cv)) + 
  tm_polygons(col="driving_SMK_20_parcel_cv", style="jenks", legend.show = FALSE) + 
  tm_shape(la_openspace) +
  tm_fill(col="green", lwd=0) +  
  tm_shape(la_water) +
  tm_fill(col="lightblue", lwd=0) + 
  tm_scale_bar(position=c("LEFT", "bottom"))
```

```{r}
cv <- ct_drive_data |>
  group_by(type, drive)

```
- repeat with ratio measures used in literature
```{r}




```

## Compare to health outcomes (need to generate evidence-based measures)
*TODO* 
- Urban vs Rural varied buffer time (e.g how the USDA does it) 
- Use ratio measures 
- Include USDA food access measures as a way to test importance 

## Map El Sendero data 
- need to link to geocoded data (ask beau for this?) 
```{r}
elsend <- read.csv("../../0_shared-data/latino-health-elsendero/raw/ElSenderoPathwaysToH_DATA_LABELS_2025-04-22_1153.csv")
elsend1 <- read.csv("../../0_shared-data/latino-health-elsendero/raw/ElSenderoPathwaysToH_DATA_LABELS_2025-04-22_1154.csv")

# map the results




```

